---
title: "Reproducible Research Project"
author: "Elbrus Gasimov, Gunneet Singh & Muhammad Usman"
date: "5/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("C:\\Users\\muhammad usman\\Desktop\\RR\\RR-Project")
```

# Introduction

This project is composed of two sections. First, it deals in reproducing original results of the study and second we will replicate same approach on another data set with some improvements which have been explained in the second section.

### Topic and Data description

This project deals in "Human Development and its effects on Economic Growth". The data set and variables used in this study spread over from 1990-2015 and includes only 5 African countries which are Kenya, Tanzania, Rwanda, Ghana and Cote d'Voire. Variables used in this study are: GDP is the dependent variable, Human Development Index (HDI) measured in percentage as the variable of interest with Inflation (INF) proxied as Consumer price index measured in percentage, Capital (CAP) proxied as Government Capital formation measured in % of GDP, Aid (AID) proxied as official development assistant measured in current USD, Investment (INV) proxied as foreign direct investment measured as (% of GDP) and Labour (LAB) as control variables. Data was retrieved from world bank and world development indicators (WDI) repositories.

## Part I Reproducibility

This part of project contains results of original study. We tried to produce original results. However we confronted with some issues in producing original results. They are discussed below step by step. Lets first load required libraries and then data set

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library("readxl")
library("lmtest")
library("MASS")
library("sandwich")
library("zoo")
library("car")
library("lmtest")
library("Formula")
library("plm")
library("stargazer")
library("aod")
```

```{r}
data <- read_excel("data.xlsx") #loading the required data set
head(data) # looking at the first 5 observations
```

Firstly, the original study took log transformations to have more normally distributed data of the following variables and then applied panel models (random and fixed effects).

```{r}
#visualizing distributions of not normally distributed features and taking log transformations
par(mfrow=c(2,3))
hist(data$INV)
hist(data$GDP)
hist(data$AID)
hist(data$LAB)

#log transformations 
data$INV<-log(data$INV)
data$GDP<-log(data$GDP)
data$AID<-log(data$AID)
data$LAB<-log(data$LAB)

#after log transformations more normally distributed data 
par(mfrow=c(2,3))
hist(data$INV)
hist(data$GDP)
hist(data$AID)
hist(data$LAB)
```

```{r}
# Estimating fixed effects model
fixed <- plm(GDP~INV+GCF+INF+AID+LAB+HDI,
            data=data, index=c("Country", "Year"), model="within")
summary(fixed)

```

We found some contrasting results in regards to original results. Variables INF, AID, LAB are significant but HDI is insignificant in our case. However, in original results all variables are significant. Furthermore, coefficients also vary in comparison to original results. Lets apply random fixed effects and see if we are able to reproduce original results.

```{r}
# Random effects model

random <-plm(GDP~INV+GCF+INF+AID+LAB+HDI, data=data, 
             index=c("Country", "Year"), model="random")
summary(random)
```

Random model throws an error that 7 coefficients to be estimated but only 5 individuals means only 5 countries. For random effects model to be estimated, we must have N \> T which means number of countries must be greater than number of years. As we cannot reproduce original results due to less number of countries as observations than time period, we would like to replicate the same study but keeping in view the requirements of panel models.

## Part II Replicability

We will now replicate the same study but on different data set and keeping in mind all the steps that are involved in performing panel models(Random and Fixed Effects). After loading the data following are the complete set of steps that we would like to perform:

1.  Checking distributions of features and taking log transformations if required

2.  Fixed effects model

3.  Random effects model

4.  Ols model

5.  Test for the selection of best model among the above mentioned

6.  Test for time and cross sectional effects

7.  General to specific approach to identify the significant variables

8.  Diagnostic tests for serial correlation and heteroskedasticity

9.  Variance-Covariance method to control both heteroskedasticity and autocorrelation

10. Conclusions

Lets first load the data set

```{r}
my_data <- read.csv("rep_data.csv")# loading csv file
head(my_data) # checking first 5 observations
```
Our data set consist of same features but spreads over from 2006 to 2016 and for total number of 14 countries. Before checking the distributions of our variables lets firts check for missing values.

```{r}
# Checking if any of the variables has NA's using sapply function
sapply(my_data, function(x) sum(is.na(x)))
```

We have no missing values so lets visualize distributions.
```{r}
par(mfrow=c(3,3)) #allows us to create a matrix of plots
par(mar=c(2,2,1,1))
hist(my_data$GDP)
hist(my_data$GCF)
hist(my_data$INF)
hist(my_data$INV)
hist(my_data$AID)
hist(my_data$HDI)
hist(my_data$LAB)
```

Apart from GCF and HDI all other distributions are either right skewed or left skewed. Lets take log transformations and see if it helps our cause. 

```{r}
#we have small values in our data set and if we perform log transformations we get negative values
#to avoid this we add positive absolute value to few of the features 
my_data$GDP<-log(my_data$GDP)
my_data$INV<-log(my_data$INV+5)
my_data$INF<-log(my_data$INF+15)
my_data$LAB<-log(my_data$LAB)
my_data$AID<-log(my_data$AID)
```

lets now check our distributions after taking log transformations.
```{r}
#visualizing log transformed features 
par(mfrow=c(3,3))
par(mar=c(2,2,1,1))
hist(my_data$GDP)
hist(my_data$GCF)
hist(my_data$INF)
hist(my_data$INV)
hist(my_data$AID)
hist(my_data$HDI)
hist(my_data$LAB)
```
The above graphs show that log transformation does help as now data looks in better shape then without log transformation. The next step is to perform panel modeling.


```{r}
# Fixed effects model
fixed <- plm(GDP~INV+INF+GCF+HDI+LAB+AID, 
            data=my_data, index=c("Country", "Year"), model="within")
summary(fixed)
```

The output of the model clearly suggests that only three out of the six variables are statistically significant at the 5% level of significance. Comparing the results with those of the original study, there is a difference in the variables which are significant. Unlike the previous dataset, HDI and GCF are significant for the current data. 

Moreover, Fixed effects model allow quantitative interpretation of the results, which means, for instance, every unit increase of HDI increases the GDP of a country by almost 10.28%


```{r}
# Creating random effects model

random <-plm(GDP~INV+INF+GCF+HDI+LAB+AID, 
            data=my_data, index=c("Country", "Year"), model="random")
summary(random)
```

Like fixed effects model, this model also suggests that only 3 and that also the same variables are significant.


```{r}
# Creating OLS model

OLS <-plm(GDP~INV+INF+GCF+HDI+LAB+AID, 
            data=my_data, index=c("Country", "Year"), model="pooling")
summary(OLS)
```

It has become almost certain that the factors "GCF", "HDI", and "LAB" are statiscally important and hence impacts GDP more than any other variables.


```{r}
# Individual effects for random effects

plmtest(OLS, type=c("bp"))
```


As the p-value is less than the 5% significance level, it means we reject the null hypothesis that all individual effects are insignificant. Clearly, it means Random effects model is better than an OLS model.

```{r}
# Individual effects for fixed effects

pFtest(fixed, OLS)
```

As the p-value is less than the 5% significance level, it means we reject the null hypothesis that all individual effects are insignificant. Clearly, it means Fixed effects model is better than an OLS model.


```{r}

```









